### 2020-03-18 notes/coherent state.tex
开始写关于相干态的note，并回顾了最基础的一些知识

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{braket}

\geometry{left=2.0cm, right=2.0cm, top=2.5cm, bottom=2.5cm}

\title{Note on coherent state}
\author{Mengzhi Wu}

\begin{document}
    \maketitle

    This note aims to summarize essential properties of the coherent state. I plan to cover these points:

    \begin{itemize}
        \item Basic propertities for creation and annilation operators.
        \item Coherent states under Hamiltonian representation.
        \item Coherent states under coordinate representation.
        \item Wigner functions for coherent states and time evolution.
        \item classical propertities of coherent states.
    \end{itemize}

    \section{Properties for operators}
        In this section, I'd like to briefly review some basic properties of some operators. The Hamiltonian is $\hat{H}=\frac{\hat{p}^2}{2m}+\frac{1}{2}m\omega^2\hat{x}^2$, and the commute relations is $[\hat{x},\hat{p}]=i\hbar$(Be careful of the sign!). Then the creation and annilation operator can be defined as

        \begin{equation}
            \hat{a}= \frac{1}{\sqrt{2m\hbar\omega}}(i\hat{p}+m\omega\hat{x}), \ \hat{a}^\dagger=\frac{1}{\sqrt{2m\hbar\omega}}(-i\hat{p}+m\omega\hat{x})
        \end{equation}

        And the commute relations is $[\hat{a}^\dagger, \hat{a}]=-1$(Be careful of the sign!). Inversely, $\hat{x}$, $\hat{p}$ and $\hat{H}$ can be described as

        \begin{equation}
            \hat{x}= ,\ \hat{p}= ,\ \hat{H}=(\hat{a}^+\hat{a}+\frac{1}{2})\hbar\omega
        \end{equation}

        Suppose the eigenstate of the Hamiltonian is $\ket{n}$ with eigenvalue $E_n=(n+\frac{1}{2})\hbar\omega$ whose wavefunction is $\Psi_n(x)=$. Then

        \begin{equation}
            \hat{a}\ket{n}=\sqrt{n}\ket{n-1},\ \hat{a}^\dagger\ket{n}=\sqrt{n+1}\ket{n+1}
        \end{equation}

        \vbox{}

        \vbox{}

        Then let's review some useful formulas. The first one is Glauber's formula, based on Baker-Hausdorff formula. This formula claims that if operators $\hat{A}$ and $\hat{B}$ satisfies $[\hat{A}, [\hat{A}, \hat{B}]] = [\hat{B}, [\hat{A}, \hat{B}]] = 0$, then

        \begin{equation}\label{formula1}
            \exp(\hat{A}+\hat{B}) = \exp(\hat{A})\exp(\hat{B})\exp\biggl(-\frac{1}{2}[\hat{A}, \hat{B}]\biggr)
        \end{equation}

        The 2nd formula is

        \begin{equation}\label{formula2}
            \exp(\lambda\hat{A})\hat{B}\exp(-\lambda\hat{A}) = \hat{B} + \lambda [\hat{A},\hat{B}] + \frac{\lambda^2}{2!} [\hat{A}, [\hat{A}, \hat{B}]] + \frac{\lambda^3}{3!}[\hat{A}, [\hat{A}, [\hat{A}, \hat{B}]]] + \dots
        \end{equation}

        Specially, if $[\hat{A}, \hat{B}]=const=:C$, then

        \begin{equation}
            \exp(\lambda\hat{A})\hat{B}\exp(-\lambda\hat{A}) = \hat{B} + \lambda C
        \end{equation}

        which means a translation of operator $\hat{B}$. More specially, we set $\hat{A}=-\alpha\hat{a}^\dagger+\alpha^*\hat{a}$, $\hat{B}=\hat{a}$ or $\hat{B}=\hat{a}^\dagger$, then we may obtain

        \begin{equation}\label{translation operator 1}
            \exp(-\alpha\hat{a}^++\alpha^*\hat{a})\hat{a}\exp(\alpha\hat{a}^+-\alpha^*\hat{a}) = \hat{a}+\alpha
        \end{equation}

        \begin{equation}
            \exp(-\alpha\hat{a}^\dagger+\alpha^*\hat{a})\hat{a}^\dagger\exp(\alpha\hat{a}^\dagger-\alpha^*\hat{a}) = \hat{a}^\dagger+\alpha^*
        \end{equation}

        Equation (\ref{translation operator 1}) is very important. Denote $\hat{D}(\alpha)=\exp(\alpha\hat{a}^\dagger -\alpha^*\hat{a})$, then (\ref{translation operator 1}) presents

        \begin{equation}\label{translation operator 2}
            \hat{D}^\dagger(\alpha)\hat{a}\hat{D}(\alpha) = \hat{a} + \alpha
        \end{equation}

        Based on (\ref{translation operator 2}), we may prove that all the eigenstates of $\hat{a}$ are

        \begin{equation}
            \hat{a} (\hat{D}(\alpha)\ket{0}) = \alpha (\hat{D}(\alpha)\ket{0})
        \end{equation}

        Therefore, we may define the coherent states as $\ket{\alpha} = \hat{D}(\alpha)\ket{0}$ with ANY complex number eigenvalue $\alpha\in\mathbb{C}$

        \vbox{}

        \vbox{}

        In the end of this section, I'd like to mention a further appliacation of (\ref{formula2}). Set $\hat{A}=\hat{N}=\hat{a}^\dagger\hat{a}$, and $\hat{B}=\hat{a}^\dagger$ or $\hat{a}$. Then we may obtain

        \begin{equation}
            \exp(\lambda\hat{N})\hat{a}\exp(-\lambda\hat{N}) = \hat{a} e^{-\lambda}
        \end{equation}

        \begin{equation}
            \exp(\lambda\hat{N})\hat{a}^\dagger\exp(-\lambda\hat{N}) = \hat{a}^\dagger e^{\lambda}
        \end{equation}

        Still pay attention to the sign! This formula will give out the squeezed states.

    \section{Coherent states under Hamiltonian representation}

    \section{Coherent states under coordinate representation}
\end{document}